---
title: "AI Integration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AI Integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE)
```

```{r, eval = TRUE}
library(TextAnalysisR)
check_ollama()
```

TextAnalysisR provides AI features via local and web-based providers.

## Providers

| Provider | Type | API Key | Best For |
|----------|------|---------|----------|
| [Ollama](https://ollama.com) | Local | None | Privacy, no cost, offline use |
| OpenAI | Web-based | OPENAI_API_KEY | Quality, speed |
| [Gemini](https://ai.google.dev/) | Web-based | GEMINI_API_KEY | Quality, speed |
| [spaCy](https://spacy.io/) | Local | None | Linguistic analysis |
| Transformers | Local | None | Embeddings, sentiment |

## Setup

### Local AI (Ollama)

```{r}
# 1. Install Ollama: https://ollama.com
# 2. Pull a model (in terminal):
#    ollama pull llama3.2
#    ollama pull mistral

# 3. Verify in R:
check_ollama()
list_ollama_models()
```

### Web-based AI (OpenAI / Gemini)

```{r}
Sys.setenv(OPENAI_API_KEY = "<openai-api-key>")
Sys.setenv(GEMINI_API_KEY = "<gemini-api-key>")

# Or use .env file in project root
# OPENAI_API_KEY=<openai-api-key>
# GEMINI_API_KEY=<gemini-api-key>
```

## Unified LLM API

```{r}
# Provider-agnostic (recommended)
response <- call_llm_api(
  provider = "openai",
  system_prompt = "You are a helpful assistant.",
  user_prompt = "Summarize this text..."
)

# Provider-specific
call_openai_chat(system_prompt, user_prompt, model = "gpt-4.1-mini")
call_gemini_chat(system_prompt, user_prompt, model = "gemini-2.5-flash")
call_ollama(prompt, model = "llama3.2")
```

## Embeddings & RAG

```{r}
# Web-based embeddings (OpenAI, Gemini)
emb <- get_api_embeddings(texts, provider = "openai")

# Local embeddings (sentence-transformers)
emb <- generate_embeddings(texts)

# RAG search over a document corpus
result <- run_rag_search(
  query = "What are the main findings?",
  documents = my_docs,
  provider = "openai"
)
```

## Ollama Utilities

| Function | Purpose |
|----------|---------|
| `check_ollama()` | Verify Ollama server is running |
| `list_ollama_models()` | List installed models |
| `get_recommended_ollama_model()` | Auto-select optimal model |

```{r}
if (check_ollama()) {
  models <- list_ollama_models()
  best_model <- get_recommended_ollama_model()
}
```

## Default Models

| Provider | Chat Model | Embedding Model |
|----------|------------|-----------------|
| OpenAI | gpt-4.1-mini | text-embedding-3-small |
| Gemini | gemini-2.5-flash | gemini-embedding-001 |
| Ollama | llama3.2 | nomic-embed-text |
| Local | - | all-MiniLM-L6-v2 |

## Responsible AI Design

All AI features follow [NIST AI Risk Management Framework](https://www.nist.gov/itl/ai-risk-management-framework) principles:

| Principle | Implementation |
|-----------|----------------|
| Human oversight | AI suggests; review and approve |
| User control | Edit, regenerate, or override any output |
| Transparency | View prompts and parameters used |
| Privacy | Local options (Ollama, spaCy) for sensitive data |
| Grounding | Content based on input data, not generic knowledge |

