---
title: "AI Integration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AI Integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE)
```

```{r, eval = TRUE}
library(TextAnalysisR)
packageVersion("TextAnalysisR")

# Local pipeline runs without any API key — TF-IDF surfaces the same
# corpus terms the LLM workflows below would summarize for you.
mydata <- SpecialEduTech[seq_len(20), c("title", "abstract")]
united <- unite_cols(mydata, listed_vars = c("title", "abstract"))
toks   <- prep_texts(united, text_field = "united_texts")
dfm    <- quanteda::dfm(toks)
extract_keywords_tfidf(dfm, top_n = 10)
```

TextAnalysisR provides AI features via cloud-based providers.

On the hosted web app, Gemini usage is free, supported by the Google Cloud Research program. OpenAI calls require a personal API key.

## Providers

| Provider | Type | API Key | Best For |
|----------|------|---------|----------|
| OpenAI | Web-based | OPENAI_API_KEY | Quality, speed |
| [Gemini](https://ai.google.dev/) | Web-based | None on hosted app; otherwise GEMINI_API_KEY | Quality, speed |
| [spaCy](https://spacy.io/) | Local | None | Linguistic analysis |
| Transformers | Local | None | Embeddings, sentiment |

## Setup

```{r}
Sys.setenv(OPENAI_API_KEY = "<openai-api-key>")
Sys.setenv(GEMINI_API_KEY = "<gemini-api-key>")

# Or use .env file in project root
# OPENAI_API_KEY=<openai-api-key>
# GEMINI_API_KEY=<gemini-api-key>
```

## Unified LLM API

```{r}
# Provider-agnostic (recommended)
response <- call_llm_api(
  provider = "openai",
  system_prompt = "You are a helpful assistant.",
  user_prompt = "Summarize this text..."
)

# Provider-specific
call_openai_chat(system_prompt, user_prompt, model = "gpt-4.1-mini")
call_gemini_chat(system_prompt, user_prompt, model = "gemini-2.5-flash")
```

## Embeddings & RAG

```{r}
# Web-based embeddings (OpenAI, Gemini)
emb <- get_api_embeddings(texts, provider = "openai")

# Local embeddings (sentence-transformers)
emb <- generate_embeddings(texts)

# RAG search over a document corpus
result <- run_rag_search(
  query = "What are the main findings?",
  documents = my_docs,
  provider = "openai"
)
```

## Default Models

| Provider | Chat Model | Embedding Model |
|----------|------------|-----------------|
| OpenAI | gpt-4.1-mini | text-embedding-3-small |
| Gemini | gemini-2.5-flash | gemini-embedding-001 |
| Local | - | all-MiniLM-L6-v2 |

## Responsible AI Design

All AI features follow [NIST AI Risk Management Framework](https://www.nist.gov/itl/ai-risk-management-framework) principles:

| Principle | Implementation |
|-----------|----------------|
| Human oversight | AI suggests; review and approve |
| User control | Edit, regenerate, or override any output |
| Transparency | View prompts and parameters used |
| Privacy | Local sentence-transformers and spaCy options for sensitive data |
| Grounding | Content based on input data, not generic knowledge |
