---
title: "Function Reference Cheatsheet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Function Reference Cheatsheet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE)
```

Quick reference guide organized by workflow stage.

## 1. Data Import & Preprocessing

| Function | Purpose |
|----------|---------|
| `import_files()` | Import CSV, XLSX, PDF, DOCX, TXT files |
| `unite_cols()` | Combine multiple text columns into one |
| `prep_texts()` | Tokenize with full preprocessing options |
| `detect_multi_words()` | Find collocations (n-grams) |
| `get_available_dfm()` | Get best available DFM with fallback |

## 2. Lexical Analysis

| Function | Purpose |
|----------|---------|
| `calculate_word_frequency()` | Count word frequencies |
| `extract_keywords_tfidf()` | TF-IDF keyword extraction |
| `extract_keywords_keyness()` | Keyness-based keywords |
| `lexical_diversity_analysis()` | TTR, MATTR, MTLD metrics |
| `calculate_text_readability()` | Flesch, SMOG, ARI scores |

<details>
<summary><strong>Visualization Functions</strong></summary>

| Function | Purpose |
|----------|---------|
| `plot_word_frequency()` | Bar chart of word frequencies |
| `plot_tfidf_keywords()` | TF-IDF keyword visualization |
| `plot_keyness_keywords()` | Keyness comparison plot |
| `plot_ngram_frequency()` | N-gram frequency plot |
| `plot_readability_distribution()` | Readability score distribution |
| `plot_lexical_diversity_distribution()` | Diversity metrics plot |

</details>

## 3. Sentiment Analysis

| Function | Purpose |
|----------|---------|
| `analyze_sentiment()` | Quick sentiment scoring |
| `sentiment_lexicon_analysis()` | Dictionary-based (no Python) |
| `sentiment_embedding_analysis()` | Neural sentiment (Python) |

<details>
<summary><strong>Visualization Functions</strong></summary>

| Function | Purpose |
|----------|---------|
| `plot_sentiment_distribution()` | Sentiment score histogram |
| `plot_sentiment_by_category()` | Sentiment by group |
| `plot_sentiment_boxplot()` | Box plot comparison |
| `plot_emotion_radar()` | Emotion radar chart |

</details>

## 4. Semantic Analysis

| Function | Purpose |
|----------|---------|
| `generate_embeddings()` | Create document embeddings |
| `reduce_dimensions()` | PCA, t-SNE, UMAP reduction |
| `calculate_document_similarity()` | Compute similarity matrix |
| `semantic_similarity_analysis()` | Full similarity workflow |
| `semantic_document_clustering()` | Cluster similar documents |
| `generate_cluster_labels()` | AI-generated cluster names |

<details>
<summary><strong>Visualization Functions</strong></summary>

| Function | Purpose |
|----------|---------|
| `plot_semantic_viz()` | 2D/3D semantic visualization |
| `plot_similarity_heatmap()` | Similarity matrix heatmap |
| `plot_cluster_terms()` | Cluster term visualization |

</details>

## 5. Network Analysis

| Function | Purpose |
|----------|---------|
| `semantic_cooccurrence_network()` | Word co-occurrence graph |
| `semantic_correlation_network()` | Word correlation graph |

## 6. Topic Modeling

| Function | Purpose |
|----------|---------|
| `find_optimal_k()` | Search for optimal topic count |
| `fit_semantic_model()` | STM (Structural Topic Model) |
| `fit_embedding_topics()` | Embedding-based topics (BERTopic) |
| `fit_hybrid_model()` | STM + embeddings hybrid |
| `get_topic_terms()` | Extract top words per topic |
| `get_topic_prevalence()` | Calculate topic prevalence |
| `generate_topic_labels()` | AI-generated topic names |

<details>
<summary><strong>Visualization Functions</strong></summary>

| Function | Purpose |
|----------|---------|
| `plot_topic_probability()` | Topic probability distribution |
| `plot_topic_effects_categorical()` | Topic effects by category |
| `plot_topic_effects_continuous()` | Topic effects over continuous var |
| `plot_word_probability()` | Word probability per topic |
| `plot_quality_metrics()` | Model quality metrics |

</details>

## 7. PDF Processing

| Function | Purpose |
|----------|---------|
| `process_pdf_unified()` | Auto-fallback PDF extraction |
| `extract_text_from_pdf()` | Extract text (R) |
| `extract_pdf_multimodal()` | Vision AI for images in PDFs |
| `detect_pdf_content_type()` | Detect PDF content type |

## 8. AI Integration

| Function | Purpose |
|----------|---------|
| `check_ollama()` | Verify Ollama availability |
| `call_ollama()` | Direct Ollama API call |
| `call_openai_chat()` | OpenAI API call |
| `generate_topic_labels_langgraph()` | Multi-agent topic labeling |
| `generate_survey_items()` | Generate survey items |

## 9. Python Environment

| Function | Purpose |
|----------|---------|
| `setup_python_env()` | Set up Python environment |
| `check_python_env()` | Check Python configuration |

## 10. Validation & Quality

<details>
<summary><strong>Validation Functions</strong></summary>

| Function | Purpose |
|----------|---------|
| `cross_analysis_validation()` | Cross-validate analysis |
| `validate_semantic_coherence()` | Check semantic coherence |
| `calculate_clustering_metrics()` | Clustering quality metrics |

</details>

---

## Launch App

The Shiny app provides an interactive interface for all functions:

```{r}
run_app()
```
