---
title: "Installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE)
```

## R Package

```{r}
install.packages("remotes")
remotes::install_github("mshin77/TextAnalysisR")
library(TextAnalysisR)
run_app()
```

**Requirements:** R >= 4.0, RStudio recommended

## Web App

Visit [textanalysisr.org](https://www.textanalysisr.org) - no installation needed.

Note: Web version has limited features (no Python, no local AI via Ollama, no large files). Cloud AI providers (OpenAI, Gemini) are available with an API key.

## Optional Features

### Linguistic Analysis (spaCy)

For lemmatization, POS tagging, and named entity recognition:

```{r}
# Python spaCy required - see setup_python_env()
TextAnalysisR::setup_python_env()
```

### Python Features

For embeddings and neural sentiment analysis:

```{r}
setup_python_env()
```

Requires Python 3.9+ and optionally [Ollama](https://ollama.com) for local AI.

## Troubleshooting

| Issue | Solution |
|-------|----------|
| Package install failed | `remotes::install_github("mshin77/TextAnalysisR", dependencies = TRUE)` |
| Browser doesn't open | Navigate to URL shown in R console |
| Python errors | See [Python Environment](python_environment.html) |

## Responsible AI Design

TextAnalysisR follows responsible AI principles with human oversight:

- **AI Suggests**: LLMs generate draft labels, content, and recommendations
- **Human Reviews**: You examine all AI outputs before use
- **Human Decides**: Edit, approve, or regenerate as needed
- **Human Controls**: Override any AI suggestion with manual input

This approach aligns with [NIST AI Risk Management Framework](https://www.nist.gov/itl/ai-risk-management-framework) and EU AI Act requirements for meaningful human control.
