---
title: "Python Environment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Python Environment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE)
```

Python enables advanced features: PDF tables, embeddings, and AI analysis.

## Quick Setup

```{r}
library(TextAnalysisR)
setup_python_env()
```

Uses conda if available, otherwise virtualenv.

## Check Status

```{r}
check_python_env()
```

## Common Issues

### "Another Python already initialized"

Set preferred environment in `.Rprofile`:

```{r}
Sys.setenv(RETICULATE_PYTHON_ENV = "textanalysisr-env")
```

Then restart R.

### Environment in OneDrive

Avoid OneDrive paths. Use:

```{r}
setup_python_env(method = "virtualenv", envpath = "C:/Python/envs")
```

## Deep Learning (Optional)

For embeddings and neural sentiment:
```bash
pip install sentence-transformers transformers torch
```

## Diagnostics

```{r}
library(reticulate)
py_config()
conda_list()
```
