---
title: "Qualitative Coding"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Qualitative Coding}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

Qualitative coding assigns codes from a codebook to units of text, then measures how far coders
agree. AI supplies suggestions; a human confirms, corrects, or rejects each one, and only reviewed
rows are exported. Nothing is coded silently.

The sections below follow the app's Qualitative Coding tabs in order. Agreement and merging run
live on constructed data. The coding steps call an AI provider, so they are shown as code rather
than run here.

```{r}
library(TextAnalysisR)
packageVersion("TextAnalysisR")
```

## 1. Codebook

A codebook is a data frame with `code` and `definition` columns, plus an optional `example`.
Definitions carry most of the weight: a code without one gives the model nothing to match against.

```{r}
codebook <- tibble::tibble(
  code = c("access", "instruction", "assessment"),
  definition = c(
    "Availability of technology, materials, or services to students",
    "Teaching methods, strategies, or curriculum delivery",
    "Measurement of student performance or progress"
  ),
  example = c(
    "tablets were provided to every student",
    "the teacher modeled each step before practice",
    "progress was measured with weekly probes"
  )
)
codebook
```

Topic labels make reasonable starting codes. The mapping is many-to-many rather than one-to-one, so
treat seeded codes as a draft to edit rather than a finished scheme.

## 2. AI coding

`apply_codes()` splits each document into units, sends each unit with the codebook, and returns one
row per unit-code pair.

```{r eval = FALSE}
texts <- SpecialEduTech$abstract[1:20]
names(texts) <- paste0("doc", seq_along(texts))

suggestions <- apply_codes(
  texts,
  codebook,
  unit = "paragraph",
  max_codes = 3,
  provider = "gemini"
)
```

Three arguments decide what the output means.

| Argument | Effect |
|----------|--------|
| `unit` | Splits on paragraphs, sentences, or whole documents. Paragraph is the usual unit of analysis. |
| `max_codes` | Allows a unit to carry none, one, or several codes. A unit that fits nothing returns `NA`. |
| `provider` | Selects OpenAI or Gemini. `auto` picks whichever key is available. |

Output columns `start` and `end` give character offsets of the unit inside its document, so a
suggestion can be traced back to the exact text it came from.

## 3. Review

Suggestions are not results. In the app each row carries a status of pending, accepted, edited, or
rejected, and only accepted and edited rows reach the export. Working from R, filter on confidence
and read the units before keeping anything.

```{r eval = FALSE}
accepted <- suggestions[!is.na(suggestions$code) & suggestions$confidence > 0.7, ]
accepted$coder <- "coder1"
```

Low-temperature settings do not guarantee stable output. `code_retest()` codes the same sample more
than once and reports how often the runs agree, next to a baseline from shuffled labels. Retest
agreement close to that baseline means the codes are not reproducible enough to analyze.

```{r eval = FALSE}
stability <- code_retest(texts, codebook, n_runs = 2, sample_n = 20)
stability$summary
```

A codebook only reaches what it was built to reach. `uncoded_units()` returns the units that
received no code, with the text sliced from the character offsets recorded during coding.

```{r}
suggestions <- tibble::tibble(
  doc_id  = c("d1", "d1", "d2"),
  unit_id = c("d1.1", "d1.2", "d2.1"),
  start   = c(1L, 30L, 1L),
  end     = c(28L, 62L, 34L),
  code    = c("access", NA_character_, NA_character_)
)

texts <- c(
  d1 = "Screen readers open the text. The course never covered any of it.",
  d2 = "Nobody explains who pays for it."
)

uncoded_units(suggestions, texts)
```

Reading this residue answers a different question from agreement: not whether coders applied the
frame consistently, but whether the frame misses content that recurs. Categories that appear
repeatedly here belong in a revised codebook or in the reporting as findings outside the original
frame.

## 4. Agreement

`code_agreement()` takes assignments from two or more coders and reports chance-corrected agreement.

```{r}
assignments <- tibble::tibble(
  doc_id = rep(paste0("doc", 1:8), each = 2),
  code   = c("access", "access", "access", "access", "access", "access",
             "access", "access", "access", "access", "access", "access",
             "access", "instruction", "instruction", "access"),
  coder  = rep(c("c1", "c2"), times = 8)
)

agreement <- code_agreement(assignments)
agreement$overall
```

Percent agreement is high here while kappa is near zero. That gap is the kappa paradox: when one
code dominates, expected agreement approaches observed agreement and kappa collapses even though
the coders rarely disagree. Gwet's AC1 and PABAK stay interpretable under that skew, which is why
all five statistics are reported together rather than kappa alone.

Disagreements are listed separately so the underlying units can be re-read.

```{r}
agreement$disagree
```

Per-code agreement shows which codes carry the disagreement.

```{r}
agreement$by_code
```

### Coders who wrote the codebook

A coder who wrote or revised the codebook has already calibrated against it, so agreement with that
coder measures shared calibration as much as shared judgement. Naming those coders in
`codebook_authors` adds an `independent` table computed among the remaining coders alone.

```{r eval = FALSE}
agreement <- code_agreement(assignments, codebook_authors = "c1")
agreement$independent
```

Report both. A gap between `overall` and `independent` indicates the frame is harder to apply
without having helped build it.

### Coders who segment differently

Grid alignment assumes coders share units. When each coder marks their own spans, boundaries rarely
match, and pivoting on unit identity would compare rows that describe different text. Coverage
alignment instead reports, for each ordered coder pair, the share of one coder's spans that overlap
a same-code span from the other.

```{r}
spans <- tibble::tibble(
  doc_id = c("doc1", "doc1", "doc1"),
  coder  = c("c1", "c1", "c2"),
  code   = c("access", "instruction", "access"),
  start  = c(1, 200, 50),
  end    = c(100, 300, 80)
)

code_agreement(spans, align = "coverage")$overall
```

Coverage is directional. One coder marking broad spans that contain another's narrow ones scores
differently in each direction, and that asymmetry is informative rather than a defect.

## Combining coder files

Each coder exports their accepted rows; `merge_codes()` binds the files into one table for
`code_agreement()`.

```{r eval = FALSE}
combined <- merge_codes(c("coder1.rds", "coder2.rds"))
code_agreement(combined)$overall
```

## Reporting

Report the unit of analysis, how many units were coded, which agreement statistic was used and why,
and the share of AI suggestions a human changed. The last figure is what separates assisted coding
from automated labeling, and it is only available when review status is recorded.
