---
title: "Qualitative Coding"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Qualitative Coding}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

Codes from a codebook are assigned to units of text, then inter-coder agreement is measured. AI
supplies suggestions; a human confirms, corrects, or rejects each one, and only reviewed rows are
exported.

Steps that call an AI provider are shown as code rather than run. The rest run on synthetic data.

```{r}
library(TextAnalysisR)
```

## 1. Codebook

A data frame with `code` and `definition`, plus an optional `example`.

```{r}
codebook <- tibble::tibble(
  code = c("access", "instruction", "assessment"),
  definition = c(
    "Availability of technology, materials, or services to students",
    "Teaching methods, strategies, or curriculum delivery",
    "Measurement of student performance or progress"
  ),
  example = c(
    "tablets were provided to every student",
    "the teacher modeled each step before practice",
    "progress was measured with weekly probes"
  )
)
codebook
```

## 2. AI coding

`apply_codes()` splits each document into units, sends each unit with the codebook, and returns one
row per unit-code pair.

```{r eval = FALSE}
texts <- SpecialEduTech$abstract[1:20]
names(texts) <- paste0("doc", seq_along(texts))

suggestions <- apply_codes(
  texts,
  codebook,
  unit = "paragraph",
  max_codes = 3,
  provider = "gemini"
)
```

| Argument | Effect |
|----------|--------|
| `unit` | Splits on sentences, paragraphs, or whole documents. Paragraph is the usual unit of analysis. |
| `max_codes` | Allows a unit to carry none, one, or several codes. A unit that fits nothing returns `NA`. |
| `provider` | Selects OpenAI or Gemini. `auto` picks whichever key is available. |

Columns `start` and `end` give character offsets of the unit inside its document.

## 3. Review

Only accepted and edited rows reach the export.

```{r eval = FALSE}
accepted <- suggestions[!is.na(suggestions$code) & suggestions$confidence > 0.7, ]
accepted$coder <- "coder1"
```

`code_retest()` codes the same sample more than once and reports how often the runs agree, next to a
baseline from shuffled labels.

```{r eval = FALSE}
stability <- code_retest(texts, codebook, n_runs = 2, sample_n = 20)
stability$summary
```

`uncoded_units()` returns the units that received no code, with the text sliced from the recorded
offsets.

```{r}
# synthetic units, paraphrased from the corpus
suggestions <- tibble::tibble(
  doc_id  = c("d1", "d1", "d2"),
  unit_id = c("d1.1", "d1.2", "d2.1"),
  start   = c(1L, 68L, 1L),
  end     = c(66L, 128L, 74L),
  code    = c("access", NA_character_, NA_character_)
)

texts <- c(
  d1 = paste("Text-to-speech tools give students access to grade-level readings.",
             "Teachers report growing confidence after the training series."),
  d2 = "The review summarizes methodological features across the included studies."
)

uncoded_units(suggestions, texts)
```

## 4. Agreement

`code_agreement()` takes assignments from two or more coders.

```{r}
# synthetic assignments
assignments <- tibble::tibble(
  doc_id = rep(paste0("doc", 1:10), each = 2),
  code   = c("access", "access", "access", "access", "access", "access",
             "instruction", "instruction", "instruction", "instruction",
             "instruction", "assessment", "assessment", "assessment",
             "assessment", "assessment", "assessment", "access",
             "access", "access"),
  coder  = rep(c("c1", "c2"), times = 10)
)

agreement <- code_agreement(assignments)
agreement$overall
```

`n` is how many units each statistic used.

| `metric` | Reads as |
|----------|----------|
| `percent` | Raw share of units both coders labeled identically. No correction for chance. |
| `pabak` | Prevalence-adjusted bias-adjusted kappa: percent agreement rescaled to the -1 to 1 range. |
| `ac1` | Gwet's agreement coefficient. Chance-corrected, stable when one code dominates. |
| `kappa` | Cohen's kappa for two coders, Fleiss' for more. Chance-corrected against observed marginals. |
| `alpha` | Krippendorff's alpha. Chance-corrected, handles missing units. |

Which units the coders differed on:

```{r}
agreement$disagree
```

Which codes carry those disagreements:

```{r}
agreement$by_code
```

Each code is scored as a yes/no indicator, so a two-code corpus returns the same figures twice, one
indicator being the other's complement.

Chance-corrected statistics turn negative when observed agreement falls below what the marginals
predict, which happens when nearly every unit carries the same code:

```{r}
# synthetic assignments, one code dominating
skewed <- tibble::tibble(
  doc_id = rep(paste0("doc", 1:8), each = 2),
  code   = c("access", "access", "access", "access", "access", "access",
             "access", "access", "access", "access", "access", "access",
             "access", "instruction", "instruction", "access"),
  coder  = rep(c("c1", "c2"), times = 8)
)

code_agreement(skewed, by_code = FALSE)$overall
```

Chance alone accounts for 0.78 against an observed 0.75, so `kappa` lands at -0.14. Only `percent`
is bounded at zero.

`codebook_authors` adds an `independent` table computed among the remaining coders alone.

```{r eval = FALSE}
code_agreement(assignments, codebook_authors = "c1")$independent
```

Everything above assumes coders shared units. When each coder highlights their own stretch of text,
`align = "coverage"` compares the highlights instead. `start` and `end` are character positions.

```{r}
# synthetic spans
spans <- tibble::tibble(
  doc_id = c("doc1", "doc1", "doc1"),
  coder  = c("c1", "c1", "c2"),
  code   = c("access", "instruction", "access"),
  start  = c(1, 200, 50),
  end    = c(100, 300, 80)
)

code_agreement(spans, align = "coverage")$overall
```

`c2`'s span sits inside `c1`'s first one. One of `c1`'s two spans was matched, so 0.50; `c2`'s only
span was matched, so 1.00. Both directions are reported.

## 5. Combining coder files

```{r eval = FALSE}
combined <- merge_codes(c("coder1.csv", "coder2.xlsx"))
code_agreement(combined)$overall
```
