---
title: "Semantic Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Semantic Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

Semantic analysis examines relationships of meaning between words and documents. The sections below
follow the Shiny app's **Semantic Analysis** tabs in order.

## Setup

A 150-document subset of `SpecialEduTech` keeps the build fast; the full dataset works the same way.

```{r}
library(TextAnalysisR)

mydata <- SpecialEduTech[1:150, ]
united_tbl <- unite_cols(mydata, listed_vars = c("title", "keyword", "abstract"))
tokens <- prep_texts(united_tbl, text_field = "united_texts", remove_stopwords = TRUE)
dfm_object <- quanteda::dfm(tokens)
```

## Word Co-occurrence

`word_co_occurrence_network()` builds a network of words that co-occur across documents, with community
detection and centrality metrics.

```{r}
network <- word_co_occurrence_network(
  dfm_object,
  co_occur_n = 30,
  top_node_n = 30,
  node_label_size = 22,
  community_method = "leiden"
)

network$plot
network$top_nodes
network$table
network$summary
```

## Word Correlation

`word_correlation_network()` connects words by phi correlation of their document co-occurrence patterns.

```{r}
corr_network <- word_correlation_network(
  dfm_object,
  common_term_n = 10,
  corr_n = 0.4,
  community_method = "leiden"
)

corr_network$plot
corr_network$top_nodes
corr_network$table
```

## Document Similarity

`semantic_similarity_analysis()` compares documents by words, n-grams, or embeddings (embeddings require
Python). The example below uses word features on a subset and renders a cosine similarity heatmap.

```{r}
subset_texts <- united_tbl$united_texts[1:10]

similarity <- semantic_similarity_analysis(
  subset_texts,
  document_feature_type = "words",
  similarity_method = "cosine",
  verbose = FALSE
)

plot_similarity_heatmap(similarity$similarity_matrix, method_name = "Cosine")
```

| Method | Description | Requires |
|--------|-------------|----------|
| Words | Word-frequency vectors (bag-of-words) | none |
| N-grams | Word-sequence vectors | none |
| Embeddings | Transformer sentence vectors | Python |

## Comparative Analysis

Comparative analysis scores how similar documents in one category are to a reference category.
`extract_cross_category_similarities()` pulls cross-category pairs from a similarity matrix and
`analyze_similarity_gaps()` summarizes the differences. The example uses the first 30 documents.

```{r}
term_matrix <- as.matrix(dfm_object)[1:30, ]
normalized <- term_matrix / sqrt(rowSums(term_matrix^2))
sim_matrix <- normalized %*% t(normalized)

docs_data <- data.frame(
  display_name = paste0("doc", 1:30),
  reference_type = quanteda::docvars(dfm_object, "reference_type")[1:30]
)
dimnames(sim_matrix) <- list(docs_data$display_name, docs_data$display_name)

cross <- extract_cross_category_similarities(
  sim_matrix,
  docs_data,
  reference_category = "journal_article",
  category_var = "reference_type",
  id_var = "display_name"
)

gaps <- analyze_similarity_gaps(cross)
gaps$summary_stats
```

## Semantic Search

`run_rag_search()` retrieves the documents most relevant to a query using embedding similarity. It
requires an OpenAI or Gemini API key; see [AI Integration](ai_integration.html).

```{r, eval = FALSE}
results <- run_rag_search(
  query = "math intervention for students with disabilities",
  documents = united_tbl$united_texts,
  provider = "openai"
)
```

## Sentiment & Emotion

`sentiment_lexicon_analysis()` scores documents with the Bing, AFINN, or NRC lexicon. NRC also yields
discrete emotions for `plot_emotion_radar()`.

```{r}
sentiment <- sentiment_lexicon_analysis(dfm_object, lexicon = "bing")
plot_sentiment_distribution(sentiment$document_sentiment)
```

```{r}
emotion <- sentiment_lexicon_analysis(dfm_object, lexicon = "nrc")
plot_emotion_radar(emotion$emotion_scores)
```

`plot_sentiment_by_category()` compares sentiment across a metadata category after joining it to the
scored documents. Transformer (`sentiment_embedding_analysis()`) and LLM (`analyze_sentiment_llm()`)
scoring require Python or an API key and are not run here.

```{r}
scored <- sentiment_lexicon_analysis(dfm_object, lexicon = "bing")$document_sentiment
scored$reference_type <- quanteda::docvars(dfm_object, "reference_type")[
  match(scored$document, quanteda::docnames(dfm_object))
]
plot_sentiment_by_category(scored, category_var = "reference_type")
```

## Document Groups

`cluster_embeddings()` groups documents from a feature matrix. K-means and hierarchical clustering run
in base R; the app's default UMAP + DBSCAN path requires Python. `generate_cluster_labels_auto()` labels
each group with its most distinctive terms.

```{r}
data_matrix <- as.matrix(dfm_object)

groups <- cluster_embeddings(data_matrix, method = "kmeans", n_clusters = 5, verbose = FALSE)
groups$n_clusters

labels <- generate_cluster_labels_auto(data_matrix, groups$clusters, method = "tfidf", n_terms = 3)
labels
```

A 2-D map of the groups uses `reduce_dimensions()` (PCA runs in base R; t-SNE and UMAP need their
packages).

```{r}
coords <- reduce_dimensions(data_matrix, method = "PCA", n_components = 2, verbose = FALSE)
head(coords$reduced_data)
```
