---
title: "Semantic Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Semantic Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE)
```

Semantic analysis finds patterns of meaning using embeddings and neural networks.

## Setup

```{r}
library(TextAnalysisR)

mydata <- SpecialEduTech
united_tbl <- unite_cols(mydata, listed_vars = c("title", "keyword", "abstract"))
tokens <- prep_texts(united_tbl, text_field = "united_texts")
dfm_object <- quanteda::dfm(tokens)
```

## Document Similarity

```{r}
similarity <- semantic_similarity_analysis(
  texts = united_tbl$united_texts,
  method = "cosine"
)
```

---

<details>
<summary><strong>Similarity Methods</strong></summary>

Semantic analysis measures document similarity using different approaches to capture meaning, from simple vocabulary matching to deep neural representations.

**Methods:**

| Method | Description | Best For |
|--------|-------------|----------|
| Words | Lexical analysis using word frequency vectors (bag-of-words) | Finding documents with shared terminology |
| N-grams | Phrase-based analysis capturing word sequences | Detecting similar phraseology |
| Embeddings | Deep semantic analysis using transformer models | Conceptual similarity, handles synonyms |

**Usage:** Choose method based on your analysis goals. Words and n-grams are faster and interpretable. Embeddings capture deeper meaning but require more computation. All methods use cosine similarity for comparison.

**Learn More:** [Sentence Transformers Documentation](https://www.sbert.net/)

</details>

---

## Sentiment Analysis

### Lexicon-based (no Python)

```{r}
sentiment <- sentiment_lexicon_analysis(dfm_object, lexicon = "afinn")
plot_sentiment_distribution(sentiment$document_sentiment)
```

### Neural (requires Python)

```{r}
sentiment <- sentiment_embedding_analysis(united_tbl$united_texts)
```

---

## Document Clustering

> **Shiny App:** Document clustering is now in **Topic Modeling → Embedding-based Topics**, which combines clustering with automatic keyword extraction.

```{r}
results <- fit_embedding_model(
  texts = united_tbl$united_texts,
  method = "umap_dbscan",
  backend = "r",
  n_topics = 5
)

results$topic_assignments
results$topic_keywords
```

For standalone clustering without keywords, use `cluster_embeddings()`. See [Topic Modeling](topic_modeling.html) for details.

### AI Cluster Labels

```{r}
labels <- generate_cluster_labels(
  results$topic_keywords,
  provider = "ollama"
)
```

---

<details>
<summary><strong>Algorithms Reference</strong></summary>

**Clustering:** K-means (spherical), Hierarchical (nested), DBSCAN (density-based), HDBSCAN (auto-detect K)

**Dimensionality Reduction:** PCA (fast, linear), t-SNE (local structure), UMAP (balanced)

</details>

---

## Network Analysis

Visualize word relationships as interactive networks with community detection.

### Word Co-occurrence Network

```{r}
network <- word_co_occurrence_network(
  dfm_object,
  co_occur_n = 10,                    # Minimum co-occurrence count
  top_node_n = 30,                    # Top nodes to display
  node_label_size = 22,               # Font size (12-40)
  community_method = "leiden"         # Community detection algorithm
)

network$plot   # Interactive visNetwork plot
network$table  # Node metrics (degree, eigenvector, community)
network$stats  # 9 network statistics
```

### Word Correlation Network

```{r}
corr_network <- word_correlation_network(
  dfm_object,
  common_term_n = 20,                 # Minimum term frequency
  corr_n = 0.4,                       # Minimum correlation threshold
  community_method = "leiden"
)
```

### Category-Specific Analysis

Enable per-category networks in the Shiny app to generate separate networks for each category, displayed in a tabbed interface.

---

<details>
<summary><strong>Network Statistics (9 Metrics)</strong></summary>

Each network returns comprehensive statistics:

| Metric | Description |
|--------|-------------|
| Nodes | Total unique words in network |
| Edges | Total connections between nodes |
| Density | Proportion of possible edges present (0-1) |
| Diameter | Longest shortest path in network |
| Global Clustering | Overall network clustering tendency |
| Avg Local Clustering | Average of local clustering coefficients |
| Modularity | Quality of community structure (higher = better separation) |
| Assortativity | Tendency of similar nodes to connect |
| Avg Path Length | Average distance between nodes |

</details>

---

<details>
<summary><strong>Community Detection Methods</strong></summary>

Community detection identifies clusters of semantically related nodes.

| Method | Description | Best For |
|--------|-------------|----------|
| `leiden` | Modern algorithm, guarantees well-connected communities | Default, best quality |
| `louvain` | Fast modularity optimization | Large networks |
| `label_prop` | Propagates labels through network | Very large networks |
| `fast_greedy` | Hierarchical agglomerative | Quick exploration |

**Learn More:** [igraph Community Detection](https://igraph.org/r/doc/communities.html)

</details>

---

## Temporal Analysis

Track themes over time:

```{r}
temporal <- temporal_semantic_analysis(
  texts = united_tbl$united_texts,
  timestamps = united_tbl$year
)
```

---

<details>
<summary><strong>Embedding Providers</strong></summary>

Document-to-document similarity using vector embeddings. Multiple providers are available:

| Provider | Model | Notes |
|----------|-------|-------|
| Ollama | nomic-embed-text (default), mxbai-embed-large, all-minilm | Free, local, private |
| OpenAI | text-embedding-3-small, text-embedding-3-large | API key required |
| Gemini | gemini-embedding-001 | API key required |
| Sentence Transformers | all-MiniLM-L6-v2 (fast), all-mpnet-base-v2 (highest quality) | Requires Python |

Embeddings are cached and shared across Document Similarity, Topic Modeling, and Semantic Search when using the same provider and model.

**Learn More:** [Sentence Transformers](https://www.sbert.net/) | [Ollama Embedding Models](https://ollama.com/blog/embedding-models)

</details>

---

## Next Steps

- [Topic Modeling](topic_modeling.html)
