---
title: "Text-Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Text-Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(TextAnalysisR)
```

The `TextAnalysisR` package provides a supporting workflow for text mining analysis. The `TextAnalysisR.app()` function allows users to launch and browse a Shiny app. This web app incorporates 'quanteda' (text preprocessing), 'stm' (structural topic modeling), 'ggraph' as well as 'widyr' (network analysis). 'tidytext' was implemented to tidy non-tidy format objects.

## Installation

The development version from [GitHub](https://github.com/mshin77/TextAnalysisR) with:

```{r, message=FALSE, eval=FALSE}
install.packages("devtools")
devtools::install_github("mshin77/TextAnalysisR")
```

## Launch and Browse the Shiny app

```{r, message=FALSE, eval=FALSE}
library(TextAnalysisR)

TextAnalysisR.app()
```

### Preprocess Text Data

```{r, message=FALSE, eval=FALSE}
data <- TextAnalysisR::SpecialEduTech 

preprocessed_data <- preprocess_texts(data, text_field = "united_texts")
```

### Plot Word Frequency

```{r, message=FALSE, eval=FALSE}
# data is a document-feature matrix (dfm) object through the quanteda package.
# Plot word frequency for the top 20 terms.

word_freq_plot <- plot_word_frequency(data, n = 20)
print(word_freq_plot)
```

### Plot Topic Per-Term Per-Topic Probabilities

```{r, message=FALSE, eval=FALSE}
# data is a tidy data frame that includes per-term per-topic probabilities (beta).
# Plot per-term per-topic probabilities for the top 10 terms.

topic_term_plot <- plot_topic_term(data, top_n = 5, ncol = 3)
print(topic_term_plot)
```

### Examine Highest Per-Term Per-Topic Probabilities

```{r, message=FALSE, eval=FALSE}
# data is a tidy data frame that includes per-term per-topic probabilities (beta).
# Examine the top 5 terms with the highest per-term per-topic probabilities.
# Number of top_n can be changed.

top_terms <- examine_top_terms(data, top_n = 5)
print(top_terms)
```

### Plot Per-Document Per-Topic Probabilities

```{r, message=FALSE, eval=FALSE}
# data is a tidy data frame that includes per-document per-topic probabilities (gamma).
# Plot per-document per-topic probabilities for the top 15 topics.
# Number of top_n can be changed.

topic_prob_plot <- topic_probability_plot(data, top_n = 15)
print(topic_prob_plot)
```

### Visualize a Table for Per-Document Per-Topic Probabilities

```{r, message=FALSE, eval=FALSE}
# data is a tidy data frame that includes per-document per-topic probabilities (gamma).
# Create a table of per-document per-topic probabilities for the top 15 topics.
# Number of top_n can be changed.

topic_prob_table <- topic_probability_table(gamma_td, top_n = 15)
print(topic_prob_table)
```




