---
title: "TextAnalysisR Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TextAnalysisR Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE
)
```

```{r, include = FALSE}
library(TextAnalysisR)
```

`TextAnalysisR` provides a supporting workflow tool for text mining analysis. The web app incorporates [quanteda](https://github.com/quanteda/quanteda) (text preprocessing), [stm](https://github.com/bstewart/stm) (structural topic modeling), [ggraph](https://github.com/thomasp85/ggraph), and [widyr](https://github.com/juliasilge/widyr) (network analysis). [tidytext](https://github.com/cran/tidytext) is implemented to tidy non-tidy format objects. The R Shiny web app is available at `TextAnalysisR::TextAnalysisR.app()` or [https://textanalysisr.org](https://textanalysisr.org). Functions are provided for completing word-topic probabilities, document-topic probabilities, estimated effects of covariates on topic prevalence, and network analysis, similar to tasks available in the web app. 

**These steps are similar to those demonstrated in the Shiny web app at `TextAnalysisR::TextAnalysisR.app()`.**

## Installation

The development version can be installed from [GitHub](https://github.com/mshin77/TextAnalysisR):

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("mshin77/TextAnalysisR")
```

## Launch and Browse the Shiny app

```{r, eval=FALSE}
library(TextAnalysisR)

if (interactive()) {
  TextAnalysisR.app()
}
```

### Unite Text Columns

```{r}
df <- TextAnalysisR::SpecialEduTech
united_tbl <- TextAnalysisR::unite_text_cols(df, listed_vars = c("title", "keyword", "abstract"))
united_tbl
```

### Preprocess Text Data

```{r}
df <- TextAnalysisR::SpecialEduTech
united_tbl <- TextAnalysisR::unite_text_cols(df, listed_vars = c("title", "keyword", "abstract"))
tokens <- TextAnalysisR::preprocess_texts(united_tbl, text_field = "united_texts", verbose = FALSE)
tokens
```

### Plot Word Frequency

```{r, fig.alt= "Word Frequency Plot"}
df <- TextAnalysisR::SpecialEduTech
united_tbl <- TextAnalysisR::unite_text_cols(df, listed_vars = c("title", "keyword", "abstract"))
tokens <- TextAnalysisR::preprocess_texts(united_tbl, text_field = "united_texts")
dfm_object <- quanteda::dfm(tokens)
word_frequency_plot <- TextAnalysisR::plot_word_frequency(dfm_object, n = 20)
word_frequency_plot
```

### Remove Common Words Across Documents

```{r, fig.alt= "Remove Common Words Across Documents"}
df <- TextAnalysisR::SpecialEduTech
united_tbl <- TextAnalysisR::unite_text_cols(df, listed_vars = c("title", "keyword", "abstract"))
tokens <- TextAnalysisR::preprocess_texts(united_tbl, text_field = "united_texts")
dfm_object <- quanteda::dfm(tokens)
dfm_removed <- TextAnalysisR::remove_common_words(tokens = tokens, 
                                   remove_vars = c("students", "study"), 
                                   dfm_object = dfm_object)
TextAnalysisR::plot_word_frequency(dfm_removed, n = 20)
```


### Evaluate Optimal Number of Topics

```{r, fig.alt= "Evaluate Optimal Number of Topics"}
df <- TextAnalysisR::SpecialEduTech
united_tbl <- TextAnalysisR::unite_text_cols(df, listed_vars = c("title", "keyword", "abstract"))
tokens <- TextAnalysisR::preprocess_texts(united_tbl, text_field = "united_texts")
dfm_object <- quanteda::dfm(tokens)
TextAnalysisR::evaluate_optimal_topic_number(
  dfm_object = dfm_object,
  topic_range = 5:30,
  max.em.its = 75,
  categorical_var = "reference_type",
  continuous_var = "year",
  height = 600,
  width = 800,
  verbose = FALSE)
```

### Plot Highest Word Probabilities for Each Topic

```{r, fig.alt= "Plot Highest Word Probabilities for Each Topic"}
df <- TextAnalysisR::SpecialEduTech
united_tbl <- TextAnalysisR::unite_text_cols(df, listed_vars = c("title", "keyword", "abstract"))
tokens <- TextAnalysisR::preprocess_texts(united_tbl, text_field = "united_texts")
dfm_object <- quanteda::dfm(tokens)
TextAnalysisR::plot_word_probabilities(
  dfm_object = dfm_object,
  topic_n = 15,
  max.em.its = 75,
  categorical_var = "reference_type",
  continuous_var = "year",
  top_term_n = 10,
  ncol = 3,
  height = 1200,
  width = 800,
  verbose = FALSE)
```

### Plot Mean Topic Prevalence Across Documents

```{r, fig.alt= "Plot Mean Topic Prevalence Across Documents"}
df <- TextAnalysisR::SpecialEduTech
united_tbl <- TextAnalysisR::unite_text_cols(df, listed_vars = c("title", "keyword", "abstract"))
tokens <- TextAnalysisR::preprocess_texts(united_tbl, text_field = "united_texts")
dfm_object <- quanteda::dfm(tokens)
TextAnalysisR::plot_mean_topic_prevalence(
  dfm_object = dfm_object,
  topic_n = 15,
  max.em.its = 75,
  categorical_var = "reference_type",
  continuous_var = "year",
  top_term_n = 10,
  top_topic_n = 15,
  height = 500,
  width = 900,
  verbose = FALSE)
```

### Plot a Word Co-occurrence Network

```{r, fig.alt= "Plot a Word Co-occurrence Network"}
df <- TextAnalysisR::SpecialEduTech
united_tbl <- TextAnalysisR::unite_text_cols(df, listed_vars = c("title", "keyword", "abstract"))
tokens <- TextAnalysisR::preprocess_texts(united_tbl, text_field = "united_texts")
dfm_object <- quanteda::dfm(tokens)
TextAnalysisR::plot_word_co_occurrence_network(
  dfm_object,
  co_occur_n = 200,
  height = 900,
  width = 800)
```

### Plot a Word Correlation Network

```{r, fig.alt= "Plot a Word Correlation Network"}
df <- TextAnalysisR::SpecialEduTech
united_tbl <- TextAnalysisR::unite_text_cols(df, listed_vars = c("title", "keyword", "abstract"))
tokens <- TextAnalysisR::preprocess_texts(united_tbl, text_field = "united_texts")
dfm_object <- quanteda::dfm(tokens)
TextAnalysisR::plot_word_correlation_network(
  dfm_object,
  co_occur_n = 30,
  corr_n = 0.4,
  height = 900,
  width = 800)
```

### Plot Word Frequency Across a Continuous Variable

```{r, fig.alt= "Plot Word Frequency Across a Continuous Variable"}
df <- TextAnalysisR::SpecialEduTech
united_tbl <- TextAnalysisR::unite_text_cols(df, listed_vars = c("title", "keyword", "abstract"))
tokens <- TextAnalysisR::preprocess_texts(united_tbl, text_field = "united_texts")
dfm_object <- quanteda::dfm(tokens)
stm_15 <- TextAnalysisR::stm_15
TextAnalysisR::word_frequency_trends(dfm_object,
                                     stm_model = stm_15,
                                     continuous_variable = "year",
                                     selected_terms = c("calculator", "computer"),
                                     height = 500,
                                     width = 900)
```
