---
title: "Topic Modeling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Topic Modeling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE)
```

Topic modeling discovers hidden themes in text collections via STM or embedding-based methods.

## Setup

```{r, eval = TRUE}
library(TextAnalysisR)

mydata <- SpecialEduTech
united_tbl <- unite_cols(mydata, listed_vars = c("title", "keyword", "abstract"))
tokens <- prep_texts(united_tbl, text_field = "united_texts")
dfm_object <- quanteda::dfm(tokens)
```

## Find Optimal Topics

```{r}
find_optimal_k(dfm_object, topic_range = 5:30)
```

## STM (Structural Topic Model)

STM discovers latent topics while incorporating document metadata as covariates.

**Best for:** metadata analysis, covariate effects, interpretable word-probability distributions.

**Quality metrics:** Semantic Coherence, Exclusivity, Held-out Likelihood.

```{r}
out <- quanteda::convert(dfm_object, to = "stm")

model <- stm::stm(
  documents = out$documents,
  vocab = out$vocab,
  K = 15,
  prevalence = ~ reference_type + s(year),
  data = out$meta
)

terms <- get_topic_terms(model, top_term_n = 10)
```

See: [Structural Topic Model](https://www.structuraltopicmodel.com/) · [stm package on CRAN](https://CRAN.R-project.org/package=stm)

## Embedding-based Topics

Uses transformer embeddings with dimensionality reduction and clustering. Best for short texts, multilingual content, and semantic similarity.

**Embedding providers:**

| Provider | Model | Notes |
|----------|-------|-------|
| Ollama (default) | nomic-embed-text, mxbai-embed-large, all-minilm | Free, local, private |
| Sentence Transformers | all-MiniLM-L6-v2, all-mpnet-base-v2 | Requires Python |
| OpenAI | text-embedding-3-small, text-embedding-3-large | API key required |
| Gemini | gemini-embedding-001 | API key required |

**Backends:** `"python"` (full BERTopic, default) · `"r"` (no BERTopic needed) · `"auto"`.

```{r}
# With precomputed embeddings
embeddings <- get_best_embeddings(
  texts = united_tbl$united_texts,
  provider = "ollama",
  model = "nomic-embed-text"
)

results <- fit_embedding_model(
  texts = united_tbl$united_texts,
  method = "umap_hdbscan",
  precomputed_embeddings = embeddings
)

# R backend (no BERTopic needed)
results <- fit_embedding_model(
  texts = united_tbl$united_texts,
  method = "umap_dbscan",
  backend = "r",
  n_topics = 10
)
```

R backend methods follow `{dimred}_{clustering}` (e.g., `"umap_dbscan"`, `"tsne_kmeans"`, `"pca_hierarchical"`).

See: [BERTopic](https://maartengr.github.io/BERTopic/) · [Sentence-BERT](https://www.sbert.net/)

## AI Topic Labels

```{r}
labels <- generate_topic_labels(terms, provider = "ollama")
```

AI generates draft labels; review and edit before final use.

## Topic-Grounded Content

Generate draft content (survey items, research questions, theme descriptions, policy recommendations, interview questions) anchored to topic term probabilities rather than generic AI knowledge.

| Content type | Output | Use case |
|------|--------|----------|
| `survey_item` | Likert-scale statement | Scale development |
| `research_question` | RQ for literature review | Systematic reviews |
| `theme_description` | Qualitative theme summary | Thematic analysis |
| `policy_recommendation` | Action-oriented statement | Policy analysis |
| `interview_question` | Open-ended question | Qualitative research |

```{r}
top_terms <- get_topic_terms(model, top_term_n = 10)

content <- generate_topic_content(
  topic_terms_df = top_terms,
  content_type = "survey_item",
  provider = "ollama",
  model = "llama3.2"
)
```

The LLM receives topic labels plus terms with beta scores. Use `get_content_type_prompt()` / `get_content_type_user_template()` to inspect or override prompts.

## Topic Prevalence Modeling

Model how topic prevalence varies with document covariates. Auto-selects Poisson, Negative Binomial, or Zero-Inflated NB based on dispersion ratio and zero proportion.

```{r}
gamma_matrix <- model$theta

prevalence_result <- fit_topic_prevalence_model(
  topic_proportions = gamma_matrix[, 1],
  metadata = out$meta,
  formula = topic_count ~ reference_type,
  model_type = "auto"
)

prevalence_result$model_type
prevalence_result$summary
prevalence_result$diagnostics
```

## STM vs Embedding

| Feature | STM | Embedding |
|---------|-----|-----------|
| Speed | Fast | Medium |
| Metadata support | Yes | No |
| Short texts | Poor | Good |
| Multilingual | No | Yes |

- **STM:** metadata analysis, covariate effects
- **Embedding:** short texts, semantic similarity (`backend = "r"` if no BERTopic)

