---
title: "Accessibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accessibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE)
```

TextAnalysisR meets WCAG 2.1 Level AA standards.

## Keyboard Navigation

| Key | Action |
|-----|--------|
| Tab / Shift+Tab | Move between elements |
| Enter / Space | Activate buttons |
| Esc | Close dialogs |

## Visual Features

- Dark mode toggle
- High contrast (4.5:1 ratio)
- 200% zoom support
- Reduced motion option

## Screen Reader Support

- Descriptive button labels
- Chart text descriptions
- Status announcements

## Language Support

- 100+ languages via Google Translate
- Auto-detected text-to-speech

## Palette Audit (developers)

Run a WCAG AA contrast audit against the app's light and dark backgrounds.
Requires the [a11yviz](https://github.com/mshin77/a11yviz) package.

```{r}
# Install once: remotes::install_github("mshin77/a11yviz")
run_app(dev = TRUE)
```

`dev = TRUE` scans the Shiny source for inline hex colors and reports any
color-background pairs that fail 4.5:1 contrast before launching the app.
