---
title: "Accessibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accessibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE)
```

```{r, eval = TRUE}
library(TextAnalysisR)

# Plots inherit accessible defaults: descriptive labels, color-blind
# safe palettes, and 4.5:1 contrast ratios verified at build time.
mydata <- SpecialEduTech[seq_len(20), c("title", "abstract")]
united <- unite_cols(mydata, listed_vars = c("title", "abstract"))
toks   <- prep_texts(united, text_field = "united_texts")
dfm    <- quanteda::dfm(toks)
plot_word_frequency(dfm, n = 10)
```

TextAnalysisR meets WCAG 2.1 Level AA standards.

## Keyboard Navigation

| Key | Action |
|-----|--------|
| Tab / Shift+Tab | Move between elements |
| Enter / Space | Activate buttons |
| Esc | Close dialogs |

## Visual Features

- Dark mode toggle
- High contrast (4.5:1 ratio)
- 200% zoom support
- Reduced motion option

## Screen Reader Support

- Descriptive button labels
- Chart text descriptions
- Status announcements

## Language Support

- 100+ languages via Google Translate
- Auto-detected text-to-speech
